Synthesizing Symmetric Lenses

### 2018-07-07 papers/SymmetricLenses/main.tex
add some edits

\section{Stochastic Regular Expressions}
\label{sec:srx}
Important to our algorithm is finding the expected number of bits needed to
recover a string in one data source from a synchronized string in the other data
source. To do this, we must first develop a probability model for our language.
We do this with \emph{stochastic regular expressions}, regular expressions
annotated with probability information. With this, we can jointly express a
language and a probability distribution over that language.

\begin{center}
  \begin{tabular}{rcl}
    \Regex{},\RegexAlt{}
    & \GEq{} & $\String$ \\
    & \GBar{} & $\emptyset$ \\
    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
  \end{tabular}
\end{center}

The language is defined like the regular expression without probability
annotations would be.  The probabilities are defined as follows:
\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

    $\ProbabilityOf{\emptyset}{\String}$
    & =
    & $0$ \\

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}
In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
how much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations.

@@

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}

\bcp{This is a POPL paper: Proofs of important theorems should
be included.  (This one might be easy enough that we can just label it as a
``Fact'' and omit the proof, or better yet we can say that the proof is straightforward.}

\subsection{Stochastic Regular Expression Entropy}
The {\em entropy} of a stochastic regular expression can be computed
directly from its syntax, when the stochastic regular expression is
unambiguous and contains only nonempty subcomponents.\bcp{What is an empty
  subcomponent?} 
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

    $\EntropyOf{\PRegexOr{\Regex}{\RegexAlt}{\Probability}}$
    & =
    & $\Probability*(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)*(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
  \end{tabular}
\end{center}
We have proven\bcp{Again, some kind of proof is needed.} that this syntactic
computation corresponds to the entropy of the stochastic regular expression.

>>>>

\section{Stochastic Regular Expressions}
\label{sec:srx}
To characterize the likely lenses, we must compute the expected number of bits
needed to recover a string in one data source from a synchronized string in the
other data source. To do this, we must first develop a probabilistic langauge
model for our language. We do this with \emph{stochastic regular expressions},
regular expressions annotated with probability information. With this, we can
jointly express a language and a probability distribution over that language.

\begin{center}
  \begin{tabular}{rcl}
    \Regex{},\RegexAlt{}
    & \GEq{} & $\String$ \\
    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
  \end{tabular}
\end{center}

In the stochastic regular expression
$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
represents the likelihood the string is in $\Regex_1$. In the stochastic regular
expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
much more likely having a string involving more than $n$ iterations is than
having a string involving $n$ iterations. The semantics of SREs is defined by
the function $P$ below.

\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String''}$
    & =
    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\

@@

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}

@@

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  The proofs of this theorem, and subsequent theorems in this paper, appears in
  the auxiliary materials accompanying this submission.
\end{proof}

\subsection{Stochastic Regular Expression Entropy}
The entropy of a data source $S$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$, the
expected number of bits required to describe a random element in $S$. The
entropy of a stochastic regular expression can be computed directly from its
syntax, when the stochastic regular expression is unambiguous.
\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{s}$
    & =
    & 1\\

@@

    $\EntropyOf{\PRegexOr{\Regex}{\RegexAlt}{\Probability}}$
    & =
    & $\Probability*(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)*(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
  \end{tabular}
\end{center}
We have proven that this syntactic computation corresponds to the entropy of the
stochastic regular expression.
